---
title: "Reproducible Research: Peer Assessment 1"
author: "Bruce Huang"
date: "Friday, February 13, 2015"
output: html_document
---
## Load Library
```{r}
library(ggplot2)
library(plyr)
library(lubridate)
library(lattice)
library(knitr)
```

## Loading and preprocessing the data
1. Load the data
2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
nike_data <- read.csv( "activity.csv", sep=",", na.strings = "NA",
                       colClasses =c("numeric","Date","numeric"))

```

## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
Daily_Step <- tapply(nike_data$steps,nike_data$date,function(x) sum(x,na.rm=TRUE))
summary(Daily_Step)
```

2. Make a histogram of the total number of steps taken each day
```{r}
qplot(Daily_Step, xlab='Number of Steps', ylab='Frequency', main="Daily Steps")
```

3. Calculate and report the mean and median of the total number of steps taken 
   per day
```{r}
mean(Daily_Step)

median(Daily_Step)
```

## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) and the average 
   number of steps taken, averaged across all days (y-axis)
```{r}
avg_pattern <- aggregate(nike_data$steps ~ nike_data$interval
                     , nike_data, FUN=mean, na.rm=T)

names(avg_pattern) <- c("interval","avg_steps")

xyplot(avg_pattern$avg_steps ~ avg_pattern$interval, 
       type = "l", ylab = "Average Steps", 
       xlab ="5-minute Interval",
       main = "Time Series Plot")
```
2. Which 5-minute interval, on average across all the days in the dataset, 
   contains the maximum number of steps?
```{r}
which.max(avg_pattern$avg_steps)
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset.
```{r}
sum(is.na(nike_data$steps))
```
2. Devise a strategy for filling in all of the missing values in the dataset.
```{r}
missing <- nike_data[is.na(nike_data),]
missing$steps <- merge(avg_pattern, missing)$avg_steps
```
3. Create a new dataset that is equal to the original dataset but with the 
   missing data filled in.
```{r}
nike_data1 <- nike_data
nike_data1[is.na(nike_data),] <- missing
Daily_Step1 <- tapply(nike_data1$steps,nike_data1$date,function(x) sum(x,na.rm=TRUE))
```
4. Make a histogram of the total number of steps taken each day and Calculate 
   and report the mean and median total number of steps taken per day.
```{r}
qplot(Daily_Step1, xlab='Number of Steps', ylab='Frequency', main="Daily Steps with missing")
mean(Daily_Step1)
median(Daily_Step1)
```

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels ¡V ¡§weekday¡¨ and
   ¡§weekend¡¨ indicating whether a given date is a weekday or weekend day.
```{r}
Weekday_type <- function(date) {
   if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {"Weekend"
        } else {"Weekday"
        }
}
nike_data1$Weekday_type <- as.factor(sapply(nike_data1$date, Weekday_type))
nike_data1$day <- sapply(nike_data1$date, FUN = Weekday_type)
```
2. Make a panel plot containing a time series plot of the 5-minute interval 
   (x-axis) and the average number of steps taken, averaged across all weekday 
   days or weekend days (y-axis). 
```{r}
avg_pattern2 <- aggregate(steps ~ interval + day, data = nike_data1, mean)
ggplot(avg_pattern2, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + 
    xlab("5-minute interval") + ylab("Number of steps")
```



